---
title: "L11- Spatial Analysis: Dependency and Interpolation "
subtitle: "AGRI4401 Precision Agriculture"
author: "Gustavo Alckmin"
format:
  revealjs:
    width: 1280
    height: 720
    margin: 0.05
    minScale: 0.5
    maxScale: 1.0
    # chalkboard: 
    #   buttons: false
    theme: [default, uwa-theme.scss] 
    #transition: slide
    css: AGRI4401_PrecisionAg.css
    embed-resources: true
    preview-links: auto
date: "2025/07/21"
date-format: long
slide-number: true
---

## Title and Welcome {background-color="darkgreen"}

### Precision Agriculture Spatial Analysis

Presented by: University Lecturer in Precision Agriculture

Today, we'll explore key concepts: Spatial Dependency, Moran's I, LISA, IDW, Kriging, and Management Zones.

These form the foundation for the R tutorial on spatial interpolation and analysis.

---

## Why Spatial Analysis in Agriculture?

In precision agriculture, fields aren't uniform. Yields, soils, and nutrients vary spatially.

Understanding spatial patterns helps optimize inputs like fertilizer and water.

We'll link theory to R code for practical mapping and zoning.

---

## Tobler's First Law

"Everything is related to everything else, but near things are more related than distant things."

This law underpins all spatial analysis in farming.

Assumption: Spatial data shows dependency, not randomness.

---

## What is Spatial Dependency?

Data like crop yields cluster: High yields near high yields due to soil, water, etc.

- Positive autocorrelation: Similar values together.
- Negative: Dissimilar neighbors.
- No autocorrelation: Random.

In R: We assess this with Moran's I on yield data.

---

## Moran's I - The Global Measure

Moran's I scores clustering from -1 to +1.

- Positive (near +1): Strong clustering – great for predictions.
- Formula compares points to neighbors, weighted by distance.

In agriculture: High Moran's I means predictable yield zones.

---

## Moran's I in Practice

Assumes spatial weights (e.g., neighbors in a grid).

In R code: We grid the yield data, compute neighbors, and run `moran.test()`.

```{r eval=F, echo=TRUE}
# Example from script: Grid and compute Moran's I
# (Simplified for presentation)
library(sf); library(spdep)
# ... (grid creation code)
moran_result <- moran.test(yield_by_cell$mean_yield, lw)
print(moran_result)
```

---

## Interpreting Moran's I Results

- High positive I: Proceed to interpolation.
- Zero: Random – interpolation risky.

In our script: Scatterplot shows quadrants (HH, LL, etc.) for visualization.

---

## Local Indicators of Spatial Association (LISA)

Moran's I is global; LISA is local.

- Identifies hotspots (HH), coldspots (LL), outliers (HL/LH).
- Significant via p-values.

---

## LISA in Agriculture

Pinpoint low-yield patches for targeted management.

- Assumes local stationarity but tests for significance.

In R: Use localmoran() and classify clusters.

```{r eval=F, echo=TRUE}
# From script: LISA computation
local_moran <- localmoran(yield_by_cell$mean_yield, weights)
# Classification and plotting
```

---

## LISA Visualization

Map shows colored clusters: Red for HH, Blue for LL.

Helps farmers spot anomalies, like a fertile spot in poor soil.

---

## Inverse Distance Weighting (IDW)

Interpolation: Estimate unsampled points from known ones.

- IDW: Closer points weigh more.
- Formula: Weighted average, weights = 1/distance^p.
- Simple, deterministic.

---

## IDW Assumptions and Use

- Assumes smooth surfaces; p tunes influence (higher p = more local).
- In agriculture: Quick yield maps.
- Limitations: No uncertainty; can create artifacts.

---

## IDW in R

From script: Use gstat::idw() with different p values.

```{r eval=F, echo=TRUE}
# Example: IDW interpolation
idw_p2 <- idw(yield ~ 1, yield_sp, grid_sp, idp = 2)
# Plot raster
# Plots compare p=1,2,3 for smoothness.
```

---

## Kriging Overview

Advanced interpolation: Best linear unbiased predictor.

- Uses variogram to model spatial structure.
- Provides predictions + uncertainty (variance).

---

## Variogram in Kriging

Plots semivariance vs. distance: Nugget, sill, range.

- Fit models like spherical.

In R: variogram() and fit.variogram().

```{r eval=F, echo=TRUE}
# From script: Variogram for meuse data
v_emp <- variogram(zinc ~ 1, meuse)
v_mod <- fit.variogram(v_emp, vgm("Sph"))
```

---

## Ordinary and Universal Kriging

- Ordinary: Constant mean.
- Universal: Accounts for trends (e.g., slope).

In agriculture: Universal for fields with gradients.

R code uses krige() with trend formulas.

---

## Co-Kriging

Incorporates secondary variables (e.g., soil moisture for yield).

- Models cross-variograms.
- Improves accuracy when variables correlate.

Script example: Zinc with distance in meuse.

```{r eval=F, echo=TRUE}
# Co-kriging code snippet
lzn.kriged = krige(log(zinc)~1, meuse, meuse.grid, model = lznr.fit)
```

---

```{mermaid}
graph TD
    A[Start: Interpolate Spatial Data in Precision Ag] --> B{Need uncertainty estimates?};
    B -- Yes --> C(Geostatistical methods);
    B -- No --> D(Deterministic methods);
    D --> E{Dataset size and complexity?};
    E -- Small/medium, simple patterns --> F[Use IDW];
    E -- Irregular points --> G[Use TIN];
    C --> I{Secondary variables available?};
    I -- Yes --> J[Use Co-Kriging];
    I -- No --> K{Spatial trends present?};
    K -- Yes --> L[Use Universal Kriging];
    K -- No --> M[Use Ordinary or Block Kriging];
```

---

## Management Zones

Divide field into uniform areas for management.

- Based on clustering interpolated data.
- Use k-means or fuzzy methods for zones.

---

## Zones in Precision Ag

- 3-5 zones typical; minimize within-zone variance.

In R: paar package with kmspc() on wheat data.

```{r eval=F, echo=TRUE}
# From script: Management zones
kmspc_results <- kmspc(wheat_sf, number_cluster = 2:4, ...)
# Plot clusters and boxplots
# Boxplots show yield by zone.
```

---

## Linking Theory to R Analysis

These concepts ground the script: Assess dependency first, then interpolate, zone.

Run the full .qmd to see outputs.

Apply to your data for real-farm insights!

Thank you! Questions?
