---
title: "L07 - Crop Spatial Variability"
subtitle: "AGRI4401 Precision Agriculture"
author: "Gustavo Alckmin"
format:
  revealjs:
    width: 1280
    height: 720
    margin: 0.05
    minScale: 0.5
    maxScale: 1.0
    # chalkboard: 
    #   buttons: false
    theme: [default, uwa-theme.scss] 
    # transition: slide
    css: AGRI4401_PrecisionAg.css
    embed-resources: true
date: "2025/07/01"
date-format: long
slide-number: true
---


## Crop Spatial Variability: An Introduction

- Spatial variability: differences in soil, crop, and landscape attributes across a field
- Example: two equal-area zones yield 1 t/ha vs. 4 t/ha (avg. 2.5 t/ha)
- Spatial pattern (clustered vs. dispersed) dictates SSCM complexity and benefit
- Soil texture (sand, silt, clay, gravel) controls nutrient storage, water-holding capacity, and chemical binding
- Soil structure (aggregate arrangement, pore distribution) influences root penetration, aeration, drainage, and erosion risk
- Multi-year monitoring under varying weather refines SSCM prescriptions

::: {.notes}
This slide introduces spatial variability, one of the two core dimensions of on-farm variability exploited by precision agriculture. Spatial variability refers to the heterogeneity of soil, crop, and landscape attributes within a single field. For example, two equal-area zones might yield 1 t/ha and 4 t/ha respectively, averaging 2.5 t/ha. These differences can be mapped with yield monitors, proximal sensing (e.g., electromagnetic surveys, gamma radiometrics), or geostatistical soil sampling.

Beyond magnitude, the spatial pattern—whether high-yield zones are clustered or dispersed—influences the complexity and economic return of site-specific crop management (SSCM). Clustered patterns facilitate management zones and variable-rate application, whereas dispersed patterns require finer control and may reduce SSCM benefits.

Soil texture is a primary driver of spatial variability. Percentage of sand, silt, clay, and gravel determines soil water retention, nutrient holding capacity, cation exchange capacity, and structural stability. Hand-textured field estimates or EM conductivity surveys guide strategic sampling to characterize these textures.

Soil structure, determined by aggregate size distribution and pore connectivity, controls root growth, water infiltration, drainage, and aeration. Compaction, excessive tillage, and organic matter depletion degrade structure, leading to poor plant-available water capacity (PAWC), oxygen stress, and root restriction. Structural impacts are highly weather-dependent (e.g., waterlogging vs. drought), so multi-season monitoring allows refinement of SSCM prescriptions and timing.
:::

---

## Basic Agronomic Principles: Overview

- Variable-rate input application driven by field variability maps
- Monitoring soil–crop–environment interactions for site-specific management
- Integration of GPS/GIS spatial data for real-time decision support
- Crop growth modeling and prescription mapping via remote and proximal sensing
- Closed-loop control of herbicide and nutrient delivery systems
- Precision sampling and geostatistical analysis to calibrate management zones

::: {.notes}
This slide introduces the foundational agronomic principles underpinning precision agriculture. Variable-rate technology (VRT) leverages field variability maps—derived from soil electrical conductivity, yield monitors, or satellite imagery—to modulate seeding, fertilization, or pesticide rates in real time (Chapters 1 and 3). Continuous monitoring of soil–crop–environment dynamics informs site-specific management, optimizing inputs while sustaining yield. GPS/GIS integration provides accurate georeferencing for operational control and decision support, enabling the creation of prescription maps using remote and proximal sensors (Chapters 2 and 8). Crop models assimilate multispectral imagery and ground truth data to simulate growth and guide application timing. Closed-loop feedback systems in irrigation and nutrient delivery automatically adjust stream rates based on sensor data, enhancing resource use efficiency (Chapter 4). Finally, precision sampling combined with geostatistical variogram analysis delineates management zones, ensuring calibration of VRT implements and accuracy of digital soil maps (Chapter 3). Together, these principles establish a data-driven framework that maximizes productivity and minimizes environmental impact.
:::

---

## Biotic Factors Affecting Variability

- Root biomass & distribution: spatial heterogeneity in root systems affects nutrient and water uptake
- Soil microbial communities: variation in bacteria, fungi, and mycorrhizae influences nutrient cycling and plant health
- Pathogen and pest pressure: localized outbreaks of diseases or insect pests create patchy yield losses
- Nematode populations: differences in nematode densities disrupt root function and crop vigor
- Organic residue decomposition: uneven breakdown rates alter soil nutrient availability
- Beneficial fauna (earthworms): variability in earthworm activity affects soil structure and fertility

::: {.notes}
Biotic factors, encompassing the living components of agroecosystems, drive significant spatial and temporal variability in crop performance. Root biomass and distribution dictate how effectively plants extract nutrients and water, with dense, well-branched root systems enhancing resource uptake in some zones. Soil microbial communities, including bacteria, fungi, and symbiotic mycorrhizae, vary across fields due to organic matter inputs, moisture gradients, and management history; these differences regulate nutrient mineralization and disease suppression. Pathogen and pest pressures often occur in hotspots, where soil conditions or crop residues favor disease proliferation, leading to patchy yield declines. Nematode populations can severely impair root health and nutrient acquisition, with populations fluctuating based on crop rotation and soil texture. The rate of organic residue decomposition varies spatially, influencing local nutrient release and immobilization patterns. Finally, beneficial soil fauna, such as earthworms, modulate soil aggregation, porosity, and nutrient cycling; their uneven distribution can create fertility gradients within a field. Understanding and mapping these biotic drivers enables targeted interventions—such as zone-specific biocontrol, root inoculants, or pest scouting—to enhance SSCM outcomes.
:::

---

## Abiotic Factors Affecting Variability

- Soil moisture variability driven by spatial heterogeneity in texture, infiltration rates, and organic matter content  
- Nutrient distribution differences (N, P, K) influenced by cation exchange capacity and historical application patterns  
- Temperature gradients across microclimates affecting crop phenology, evapotranspiration, and stress tolerance  
- Rainfall intensity and distribution creating zones of waterlogging or drought stress within a single field  
- Soil structure heterogeneity (compaction, porosity) altering root proliferation and hydraulic conductivity  
- Electrical conductivity mapping to infer salinity, moisture status, and subsurface layering  

::: {.notes}
Abiotic variability arises from dynamic interactions between physical and chemical soil properties and local weather patterns. Spatial differences in texture and organic matter drive moisture retention variability, while cation exchange capacity governs nutrient availability and leaching risk. Microclimatic temperature variation influences evapotranspiration rates and crop development stages, often creating stress hotspots. Precipitation patterns further modulate these gradients by inducing localized waterlogging or deficit zones. Soil compaction and porosity heterogeneity restrict root growth and alter hydraulic pathways, exacerbating moisture stress. Precision tools such as on-the-go soil electrical conductivity sensors (Adamchuk et al., 2004; USDA-ARS) enable high-resolution mapping of these abiotic factors, feeding variable-rate application systems to optimize input placement and mitigate yield-limiting variability.
:::

---

## Management Practices and Variability

- Integrate multiple geospatial and statistical layers via clustering to define management zones without arbitrary boundaries  
- Validate input maps to remove artifacts from mismanagement or data-collection errors and ensure true field variability  
- Recognize diminishing inter-class mean differences as zone count increases, reducing agronomic and economic justification for excessive classes  
- Balance statistical distinctiveness with agronomic significance when selecting the optimal number of zones  
- Choose appropriate management scale—within-field vs. between-field/regional—based on variability patterns and mapping-unit definitions  
- Align zone selection with economic returns, input cost reduction, and environmental objectives  

::: {.notes}
This slide discusses the critical considerations for delineating and managing precision agriculture zones. By clustering multiple geospatial and statistical data layers, we create management zones that reflect true field heterogeneity, avoiding arbitrary splits. It is essential to pre-process input maps to remove artifacts from past mismanagement or sensor errors, ensuring that zones represent genuine variability. As more zones are defined (e.g., comparing classes in Figures 7.5e,f with 7.6e,f), the mean-value differences between zones diminish, which can undermine both agronomic benefits and return on investment. Therefore, determining the optimal number of zones requires balancing statistically significant distinctions against practical management constraints. Additionally, decision-makers must select the proper scale of variability—within-field or between-field/regional—depending on whether the goal is fine-scale input optimization or broader landscape-level planning. The chosen scale should maximize economic returns, reduce input costs, and support environmental sustainability, recognizing that management strategies will differ across scales.
:::

---

## Biochemical & Biophysical Properties Soil: Overview

- Apparent soil electrical conductivity (ECa) integrates soil moisture, soluble salts, clay content, and temperature influences.
- Measurement via electromagnetic induction (horizontal/vertical coil), electrical resistivity, and time-domain reflectometry.
- Profiling delineates spatial variability of edaphic properties: clay layers, flood-deposited sands, and organic matter gradients.
- Applications include salinity mapping, assessment of irrigation/drainage efficiency, and detection of compaction zones.
- Strong correlations between ECa and crop yield inform site-specific management and variable-rate interventions.
- Georeferenced ECa mapping through GPS/GIS integration enables precision agriculture decision support.

::: {.notes}
Apparent soil electrical conductivity (ECa) is a non-invasive proxy measurement reflecting multiple soil attributes such as water content, soluble salts, clay mineralogy, bulk density, and temperature. Initially developed for salinity mapping in arid, irrigated lands, ECa profiling has evolved to capture fine-scale spatial heterogeneity in soil texture, moisture retention zones, organic matter distribution, and anthropogenic soil disturbances. Key measurement technologies include electromagnetic induction in horizontal and vertical coil orientations, electrical resistivity tomography, and time-domain reflectometry, each offering distinct penetration depths and resolution trade-offs. When coupled with high-precision GPS and GIS platforms, ECa datasets can be interpolated to generate continuous maps that correlate strongly with crop yield variability. These insights inform variable-rate irrigation, fertilization, and tillage prescriptions. Emerging research focuses on standardizing calibration protocols, integrating sensor fusion approaches, and leveraging machine learning to enhance ECa-based predictive models for sustainable site-specific management.
:::

---

## Leaf Area Index (LAI)

- Definition: one-sided green leaf area per unit ground area (m²/m²).
- Measurement methods: hemispherical photography, LiDAR, multispectral and hyperspectral sensors.
- Remote-sensing integration: NDVI- and radar-based LAI retrieval at field to landscape scales.
- Crop monitoring: LAI as proxy for biomass, growth stage, and photosynthetic capacity.
- Decision-making: informing tactical and operational adjustments to N fertilization and irrigation rates.
- Functional units: combining LAI maps with soil data for site-specific input application.

::: {.notes}
Leaf Area Index (LAI) is a key biophysical parameter in precision agriculture, quantifying the ratio of leaf surface area to ground area. According to Stoorvogel et al., LAI can be measured in situ using hemispherical photography and LiDAR or estimated remotely via multispectral and radar sensors. NDVI-based algorithms translate spectral reflectance into LAI values, providing spatially continuous maps at resolutions from sub-meter (UAV) to 10 m or coarser (satellite). LAI correlates strongly with canopy biomass and photosynthetic capacity, making it a robust indicator for crop growth stage and health. In the tactical layer, LAI time-series guide seasonal plan updates, while in the operational layer, real-time or near-real-time LAI maps enable variable-rate nitrogen and water applications. Integrating LAI with digital soil maps and yield data yields functional management units, optimizing input use efficiency and sustaining soil health.
:::

---

## Chlorophyll & Anthocyanin Content

- Chlorophyll concentration estimated via red-edge and near-infrared reflectance indices (e.g., CI, NDRE) correlates with leaf nitrogen status
- Anthocyanin pigments detected using green-red and blue-green spectral bands (e.g., ARI1, ARI2, GRVI) as stress and maturation indicators
- Proximal sensors (SPAD meters) and remote platforms (multispectral/hyperspectral UAVs) capture reflectance at key wavelengths (520–550 nm, 660–700 nm, 750–950 nm)
- Radiative transfer models (e.g., PROSPECT) and empirical regressions translate canopy spectra into pigment concentration (μg cm⁻²)
- High‐resolution maps of pigment distribution enable variable‐rate N fertilization, early stress detection, and yield forecasting
- Integration with GIS and yield monitors supports spatio-temporal decision-making for site‐specific crop management

::: {.notes}
Chlorophyll and anthocyanin are two critical biochemical markers used to assess crop health and stress. Chlorophyll absorbs strongly in the red (660–680 nm) and exhibits a sharp reflectance increase at the red-edge (700 nm). Indices such as the Chlorophyll Index (CI) and Normalized Difference Red Edge (NDRE) exploit these spectral features to estimate leaf chlorophyll concentration, which is closely linked to nitrogen status and photosynthetic capacity. Anthocyanins, secondary metabolites, absorb in the green (520–550 nm) and blue regions; indices like ARI1/ARI2 and the Green-Red Vegetation Index (GRVI) quantify their levels, revealing stress responses from drought, nutrient deficiency, or maturation. Leaf-level sensors (SPAD) provide instantaneous, nondestructive readings, while multispectral and hyperspectral platforms on UAVs or fixed wings generate georeferenced maps. Radiative transfer models like PROSPECT, calibrated with laboratory and field measurements, convert canopy reflectance into absolute pigment concentrations (μg cm⁻²). These data feed into GIS layers and variable-rate applicators to optimize nitrogen management, detect early onset of stress, support phenotyping trials, and improve yield predictions in precision agriculture.
:::

---

## Solar Induced Fluorescence

- Solar Induced Fluorescence (SIF) is chlorophyll’s re-emission of absorbed photons in the 650–800 nm spectral range in vivo
- Strongly correlates with photosynthetic electron transport and gross primary production (GPP)
- Acquired by high-resolution spectrometers on satellites, UAVs, or tower-based platforms
- Offers real-time, direct proxy for canopy light use efficiency, less confounded by external reflectance
- Sensitive to plant stress (nutrient, water, disease) enabling targeted management
- Integrates over the canopy, supporting variable rate irrigation and fertilization decisions

::: {.notes}
Solar Induced Fluorescence (SIF) emerges when chlorophyll molecules absorb sunlight and re-emit a fraction of that energy at longer wavelengths (650–800 nm). Unlike passive vegetation indices, SIF directly traces the photochemical processes of photosynthesis, providing an instantaneous proxy for gross primary production (GPP) and light use efficiency. Research has demonstrated strong linear relationships between SIF signals and measured carbon assimilation across crop species under varying stress conditions.

Advances in spectrometer technology allow retrieval of sub-nanometer SIF peaks from platforms ranging from low-flying drones to orbiting satellites. Tower-based systems can capture diurnal dynamics, while satellite missions like ESA’s FLEX will soon deliver global continuous SIF maps. In precision agriculture, this information is pivotal: spatial and temporal SIF patterns can reveal zones of water or nutrient limitations, early disease onset, or suboptimal canopy development, guiding variable rate irrigation and fertilization to optimize yield and resource use efficiency.
:::

---

## Grain Quality: Protein, Oil, Starch

- Protein variability mapping using NIR sensors to detect near-infrared reflectance for real-time quality assessment
- Mid-season NDVI and chlorophyll sensors to guide in-season N rates for targeted protein enhancement
- Harvest-derived protein maps integrated into variable-rate N prescriptions for site-specific quality management
- Zone-based VR fertiliser programs aligning the Six Rs framework to stabilise protein, oil and starch concentrations
- Sectional control on spreaders to reduce overlap and edge losses, ensuring uniform grain composition
- Long-term VR N strategies adapting to soil trends for consistent grain quality across seasons

::: {.notes}
Management of grain quality parameters—protein, oil and starch—is a key objective in modern precision agriculture. Combine-mounted NIR sensors produce high-resolution protein maps at harvest, which research shows can drive protein-based variable-rate N prescriptions to improve uniformity in grain composition. In-crop canopy sensors (e.g., NDVI, SPAD) allow mid-season adjustments of N rates, optimising protein accumulation without sacrificing yield potential. Implementing the Six Rs (right source, rate, time, place, etc.) within VR fertiliser programs ensures that nutrient supply matches crop demand for consistent quality. Sectional control technology on spreaders and sprayers eliminates overlaps and edge losses, further homogenising grain attributes across the field. Multi-year VR N strategies, informed by soil N dynamics and yield trends, provide stability in protein and starch content under variable seasonal conditions.
:::

---

## Monitoring Crop Properties: Overview

- Spectral reflectance indices (e.g., NDVI, red-edge) for chlorophyll and biomass estimation
- Canopy thermal imaging to detect plant water stress and transpiration dynamics
- In-situ soil moisture sensing (TDR, capacitance probes) synchronized with crop uptake patterns
- UAV and satellite-based multispectral/hyperspectral imaging for nutrient deficiency mapping
- Proximal sensors (chlorophyll meters, fluorescence) for real-time leaf health assessment
- Integration into DSS for variable-rate irrigation and fertilization prescriptions

::: {.notes}
Precision agriculture leverages multiple sensing modalities to quantify spatial variability in crop properties. Spectral reflectance indices such as NDVI and red-edge wavelengths correlate strongly with leaf chlorophyll content, leaf area index, and biomass, providing non-destructive estimates of plant health. Thermal imaging captures canopy temperature differences caused by stomatal conductance and water stress, enabling early detection of irrigation needs. In-situ soil moisture sensors (TDR and capacitance) provide high-temporal-resolution data on root-zone water availability, which can be aligned with crop water uptake models. UAV and satellite-based multispectral or hyperspectral platforms can cover large areas rapidly, identifying localized nutrient deficiencies via narrowband reflectance signatures. Proximal sensing tools like handheld chlorophyll meters and fluorometers deliver point measurements for calibration and validation of remote sensing models. All sensor data streams feed into decision support systems (DSS) to generate variable-rate irrigation and fertilization maps, optimizing input application only where and when the crop requires it, thereby enhancing resource-use efficiency and crop productivity while reducing environmental impacts.
:::

---

## Remote Sensing of LAI

:::: {.columns}


::: {.column}
- Definition and importance of Leaf Area Index (LAI) in crop growth modeling
- Satellite platforms (Sentinel-2, Landsat 8) and UAV-borne multispectral sensors for LAI mapping
- Spectral indices (NDVI, EVI, Red-Edge) and narrow-band algorithms tailored to canopy structure
- Retrieval methods: empirical regression models vs. physical radiative transfer model inversion
- Spatial resolution (10–30 m vs. sub-meter) and temporal revisit trade-offs for monitoring dynamics
- Integration of LAI products into precision management: stress detection, variable-rate irrigation and fertilization
:::

::: {.column}

```{mermaid, }
graph TD
  A[Data Acquisition: Satellite] --> B[Preprocessing: Atmospheric Correction]
  B --> C[Compute Spectral Indices NDVI, Red-Edge]
  C --> D{Retrieval Method}
  D -->|Empirical| E[Regression Model]
  D -->|Physical| F[RTM Inversion]
  E --> G[LAI Map]
  F --> G[LAI Map]
  G --> H[Precision Application: Scouting, VRF]
```
:::

::::


::: {.notes}
Leaf Area Index (LAI) quantifies green leaf area per ground unit and is a key biophysical variable for estimating canopy photosynthesis, evapotranspiration, and biomass. Remote sensing platforms provide synoptic, repeatable observations: Sentinel-2 offers 10 m multispectral bands including red-edge, while UAVs can capture sub-meter resolution data crucial for smallholder plots and heterogeneous fields. Spectral indices such as NDVI and Enhanced Vegetation Index (EVI) are correlated with LAI but saturate at high leaf densities; red-edge indices and narrow-band retrievals mitigate saturation effects.

Two main retrieval approaches exist: empirical regressions calibrated with ground truth LAI measurements (e.g., ceptometer, hemispherical photography) and physically based inversions of radiative transfer models (e.g., PROSAIL) that account for leaf optical properties and canopy structure. Users must balance spatial resolution against revisit frequency: satellites provide regular coverage for seasonal trends, while UAVs deliver fine-scale detail for operational adjustments.

LAI maps feed into decision-support systems for variable-rate irrigation and fertilization, enabling site-specific resource optimization and early stress detection.
:::

---


## Grain Quality Sensors In-Field

- On-combine near-infrared (NIR) sensors for real-time grain protein and moisture measurement
- GPS-referenced data capture enabling spatial protein mapping across paddocks
- Development of variable-rate N application prescriptions from protein maps to target low-protein zones
- Integration with yield monitors and GIS for overlaying protein, yield, and soil data layers
- Utilisation of 4R framework in sensor-driven protein management to optimise N source, rate, timing, and placement
- Adaptive in-season adjustments reducing N in waterlogged or high-protein areas to mitigate environmental risk

::: {.notes}
Real-time near-infrared (NIR) spectroscopy mounted on combines provides continuous measurement of grain protein concentration (GPC) and moisture at harvest. Peer-reviewed studies by CSIRO and university groups validate sensor accuracy within ±0.2% GPC under field conditions. Geographic coordinates tagged to each sample enable the creation of high-resolution protein maps when combined with yield monitor outputs and soil characteristics. These maps support precision nitrogen management through the generation of variable-rate (VR) prescriptions that specifically target underperforming or low-protein zones.

Applying the 4R nutrient stewardship framework, these technologies refine the right source, rate, time, and place for in-season N adjustments. Research trials documented in PA in Practice III (2024) demonstrate yield and protein gains of 5–10% when VR-N is guided by protein mapping, while also reducing environmental nitrogen losses in areas prone to waterlogging. The system’s adaptability allows farmers to reduce urea applications in poorly drained zones and maintain optimal protein in well-drained areas, thus balancing agronomic performance with sustainability objectives.

Ongoing case studies across mixed soil landscapes and diverse crop rotations highlight stepwise adoption pathways. Initial calibration on small paddocks, followed by full-scale integration, has proven effective in delivering economic returns of up to $50/ha and lowering N surplus by 15–25 kg N/ha. These outcomes underscore the role of in-field grain quality sensors in enhancing profitability and environmental stewardship within modern precision agriculture.
:::

---

## Spectral Data & Vegetation Interaction

- Calculation of key vegetation indices (NDVI, NDRE, PRI) from multispectral/UAV sensors
- Correlation of spectral reflectance with canopy biomass and chlorophyll content
- Spatiotemporal analysis of crop stress and water uptake dynamics
- Calibration using ground truth from neutron-probe and soil-core volumetric data

::: {.notes}
This slide explores how spectral remote sensing integrates with EM-based soil water mapping to enhance within-field crop diagnostics. We start by calculating standard vegetation indices—NDVI for biomass, NDRE for chlorophyll, and PRI for light-use efficiency—using UAV or tractor-mounted multispectral cameras. These indices are empirically correlated to canopy water status and biomass yield. By overlaying index maps with depth-resolved EM conductivity layers and ground-truth moisture measurements (neutron probes and soil cores), we can refine regression models that predict real-time soil-plant water relations. The resulting datasets enable dynamic monitoring of crop stress zones, informing variable-rate irrigation and nitrogen application. This approach has been validated in barley trials, showing improved water-use efficiency and yield uniformity across contrasting soil textures.
:::

---


## Role of Yield Monitors in Precision Ag

- Real-time mass flow and moisture sensors capture instantaneous yield data
- GPS-stamped yield maps enable high-resolution spatial analysis of productivity
- Calibration protocols ensure sensor accuracy across varying crop conditions
- Integration with prescription files allows dynamic variable-rate nutrient application
- Temporal yield trends facilitate delineation of management zones and decision making
- Continuous feedback loop supports long-term optimization of input use and ROI

::: {.notes}
Yield monitors mount on harvesters to continuously record grain flow and moisture, producing spatially referenced yield maps. Accurate calibration against load cells and moisture standards is critical to minimize systematic bias. Once validated, yield data integrate with prescription maps to drive variable-rate fertiliser or seeding operations, enabling within-paddock adjustments in response to actual performance. Temporal analysis of multi-year yield layers reveals stable management zones, informing nutrient stewardship under the “6 Rs” framework. Real-world case studies demonstrate 5–12 % ROI improvements when yield monitor data guide VR-N programs. Ongoing sensor maintenance and data validation ensure reliability, closing the loop between monitoring, decision support, and agronomic outcomes.
:::

---

## Importance of Yield Map Interpretation

- Spatial vs temporal variability in yield patterns
- Impact of variability pattern (patch uniformity vs mosaic complexity)
- Measurement scales: within-field mapping to whole-farm surveys; timing across growth stages/seasons
- Role of soil texture: sand/silt/clay distribution; EC and gamma-radiometric sensing for texture zoning
- Influence of soil structure: aggregation, porosity, degradation impacts on root growth, water availability, and erosion
- Guiding SSCM through yield map interpretation to optimize input allocation

```{mermaid}
graph LR
    A[Yield Data Collection] --> B[Yield Map Analysis]
    B --> C[Variability Pattern Identification]
    C --> D[Management Zone Delineation]
    D --> E[Site-Specific Input Application]
```

::: {.notes}
Yield maps are foundational to SSCM by revealing both spatial and temporal variability within fields. Spatial variability arises from inherent differences in soil, topography, and microclimate, while temporal changes reflect seasonal or year-to-year shifts due to weather, management, or soil health dynamics. The pattern of variation—whether in large uniform patches or fine-scale mosaics—dictates management complexity and potential ROI for variable-rate technologies. Accurate interpretation requires multi-scale measurement: high-resolution yield monitoring during harvest, complemented by whole-field surveys across seasons and growth stages. Soil texture mapping via EC and gamma-radiometric sensors further refines zones by identifying sand, silt, and clay proportions that affect water- and nutrient-holding capacities, whereas assessments of soil structure—aggregate stability, pore distribution, and compaction—predict root penetration and drainage behavior. Integrating these data layers enables delineation of management zones tailored to local limitations, optimizing fertilizer, irrigation, and cultivar choices for enhanced productivity and resource efficiency.
:::

---

## Yield Map Data Cleaning

- Seven years of georeferenced yield data (wheat, maize, sunflower, sorghum) aggregated
- Spatial structure quantified through semi-variogram fitting for each dataset
- Simple kriging interpolation to a unified 30 m grid across all years
- Standardization of yield values and detection of outliers via statistical thresholds
- Alignment and stacking of annual yields into a multispectral yield cube
- Quality-control checks to ensure spatial and temporal consistency

::: {.notes}
In this slide, we detail the preprocessing steps applied to seven years of georeferenced yield observations across four crops. First, each annual yield dataset underwent semi-variogram fitting to model spatial autocorrelation and quantify nugget, sill, and range parameters. These variogram models then drove simple kriging interpolation, producing continuous yield surfaces at 30 m resolution for each year and crop. Following interpolation, yield values were standardized (z-scores) to normalize across years and crops, and robust outlier detection was performed using a 3-sigma threshold. The cleaned, standardized yield layers were then aligned spatially and temporally to form a multispectral yield cube. Final quality-control metrics, such as pixel-wise variance checks and cross-year correlation analyses, verified that the cleaned data maintain high spatial integrity and are suitable for subsequent fuzzy c-means clustering into management zones.
:::

---

## Calibrating Yield Monitors

- Pre-season sensor inspection: clean sensors, inspect cables and slip rings
- Mass flow calibration: collect reference weights across multiple loads to establish sensor-voltage curve
- Ground speed validation: compare GNSS-derived speed with radar/encoder readings
- Moisture sensor tuning: perform gravimetric tests to derive moisture correction coefficients
- System parameter configuration: input header width, combine settings, and sensor-specific parameters
- Calibration record-keeping: document settings by crop type, speed, and moisture conditions

::: {.notes}
Accurate yield monitoring is foundational for precision agriculture analytics. Begin with a thorough pre-season inspection: residues or damaged cables can introduce noise into optical or impact-based sensors. For mass flow calibration, collect multiple grain samples of known weight on the head and compare against the monitor’s raw voltage output, fitting a calibration curve (often linear regression). Ground speed errors directly translate to area calculation errors—verify GNSS speed against a radar or shaft encoder, adjusting parameters if discrepancies exceed 1%. Moisture content affects weight; perform gravimetric moisture tests in the lab or with a handheld meter to calculate moisture correction coefficients for the on-board sensor. Ensure all system parameters (header width, combine fan speed, cleaning auger settings) are accurately entered into the monitor software. Finally, maintain detailed records of calibration settings across different crops, operating speeds, and moisture conditions. This systematic approach minimizes measurement bias and ensures high-fidelity yield maps for downstream prescription development.
:::

---

## Using Crop Quality Sensors

::::{.columns}

::: {.column}

- Sensor types: optical imaging, NIR spectroscopy, chlorophyll fluorescence
- Real-time indices: NDVI, PRI, anthocyanin content, moisture levels
- Integration with UAVs and ground rovers for high spatial resolution
- Calibration protocols and chemometric modeling for quantitative accuracy
- Data fusion: combining spectral, thermal and morphological measurements
- Decision support: linking sensor outputs to quality grading and harvest timing

:::
::: {.column}


```{mermaid}
graph TD
  A[Crop Quality Sensors] --> B[Data Acquisition]
  B --> C{Data Processing}
  C --> D[Quality Indices]
  C --> E[Spatial Mapping]
  E --> F[Zone-Based Decisions]
  D --> G[Quality Grading & Harvest Timing]
```
:::

::::

::: {.notes}
Crop quality sensors enable non-destructive, high-throughput assessment of key quality attributes such as sugar content, moisture, pigment concentration and texture. Optical imaging (multispectral, hyperspectral) and NIR spectroscopy capture spectral signatures that correlate with biochemical constituents. Fluorescence methods (e.g., chlorophyll fluorescence) assess plant stress and maturity. Calibration against laboratory reference data and chemometric techniques (PLS regression, machine learning algorithms) are essential to translate raw sensor signals into accurate quality indices. Platforms include UAVs for rapid aerial surveys and ground rovers for localized scanning. Data fusion integrates multiple modalities (spectral, thermal, structural) to enhance predictive power. Processed outputs feed decision support systems that classify produce into quality grades and optimize harvest schedules, minimizing waste and maximizing market value. Ongoing research focuses on miniaturization, real-time analytics and scalable deployment in large-scale mechanized farms.
:::

---

## Analyzing Yield Maps

- Yield maps visualize spatial yield variability across fields (t/ha).
- Geostatistical tools (variogram, kriging) identify high- and low-yield zones.
- Temporal trend analysis monitors per-grid stability over multiple seasons.
- Variability metrics (CV, range) quantify magnitude and consistency.
- Overlay yield zones with soil texture and structure data for correlation.
- Generate management zones to guide variable-rate input applications.

::: {.notes}
Yield maps are foundational in precision agriculture, capturing harvester-collected t/ha data georeferenced across the field. Spatial interpolation techniques like variogram modeling and kriging transform discrete point data into continuous yield surfaces, revealing yield patterns and gradients. The variogram quantifies spatial dependence, informing optimal interpolation parameters. Temporal analysis, comparing yield surfaces across seasons, highlights trend stability and identifies persistent zones of under- or over-performance. Statistical measures such as coefficient of variation and yield range provide insights into the magnitude of variability and its economic significance. Integrating yield patterns with soil attribute maps (texture from EC or gamma-radiometry, structure from bulk density surveys) enables causal inference and refines management prescriptions. Resulting management zones support site-specific variable-rate applications, enhancing input efficiency and yield potential.
:::

---

# Spectral Data

---

## Irradiance & Radiance

```{mermaid}
flowchart LR
    Sun["Sun: E(lambda) Irradiance (W/m2 nm)"] -->|Incident Flux| Target["Canopy/Soil"]
    Target -->|Reflected/Emitted Radiance| Sensor["Optical Sensor"]
```

- Irradiance (E(λ)): spectral power per unit area incident normal to the surface (W·m⁻²·nm⁻¹).
- Radiance (L(λ,θ,φ)): spectral power per unit area per unit solid angle emitted or reflected (W·m⁻²·sr⁻¹·nm⁻¹).
- Bidirectional dependence: Radiance varies with sensor viewing angle (θ) and azimuth (φ).
- Sensor measurement integrates radiance over its field-of-view and spectral bandwidth.
- Radiance retrieval: Reflectance(λ,θ,φ) * Irradiance(λ) / π for Lambertian surfaces.
- Crucial for calibrating vegetation indices and quantitative modeling.

::: {.notes}
This slide introduces the fundamental radiometric quantities used in optical remote sensing: irradiance and radiance. Irradiance E(λ) defines the spectral flux density incident upon a surface, whereas radiance L(λ,θ,φ) quantifies the directional spectral flux density leaving the surface per unit solid angle. The diagram depicts the sun emitting irradiance onto the canopy or soil, which then reflects or emits radiance measured by a sensor. The bidirectional nature of radiance requires careful consideration of sensor viewing geometry (θ, φ) and field-of-view. For Lambertian approximations, radiance can be related to surface reflectance and incident irradiance divided by π. Accurate quantification of these measures is critical for calibrating vegetation indices, energy balance models, and retrieving biophysical parameters in precision agriculture.
:::

---

## Reflectance, Transmittance & Absorption

- Definition of reflectance (ρ): ratio of reflected to incident radiation, wavelength dependent
- Transmittance (τ): fraction of radiation passing through a medium, influenced by thickness and optical properties
- Absorption (α): portion of energy retained by soil or plant tissues, driving photochemical processes and heat conversion
- Spectral signatures: basis for remote sensing differentiation of crop health and soil composition
- Beer–Lambert Law: quantifies absorption relative to path length and absorber concentration
- Energy conservation: ρ + τ + α = 1, foundational for radiative transfer models in precision agriculture

::: {.notes}
In precision agriculture, understanding how solar radiation interacts with soil and vegetation is fundamental for remote sensing and crop monitoring. Reflectance (ρ) varies with wavelength, surface roughness, and material properties, providing spectral signatures that distinguish crop species, stress levels, and soil characteristics. Transmittance (τ) governs the fraction of radiation that penetrates leaves and soil layers, critical for canopy penetration models and subsurface sensing. Absorption (α) drives photosynthesis in plant tissues and influences soil thermal regimes. The Beer–Lambert law formalizes the relationship between absorber concentration, path length, and absorption, underpinning quantitative retrievals of chlorophyll content and soil moisture. Conservation of energy (ρ + τ + α = 1) is central to radiative transfer models such as PROSAIL, enabling inversion of biophysical parameters from multispectral and hyperspectral data. Accurate partitioning of radiation supports calculation of vegetation indices (e.g., NDVI, PRI) and informs site-specific management practices.
:::

---

## Applying Spectral Principles to Crops

- Plant reflectance signatures across VIS, NIR and SWIR wavelengths
- Common vegetation indices (NDVI, PRI, EVI) for monitoring physiology
- Hyperspectral potential detection of nutrient deficiencies and water stress
- Sensor calibration and radiometric correction workflows
- Generation of variable-rate application prescriptions from spectral maps

```{mermaid}
graph LR
  UV[UV 380nm ] --> Visible[Visible 400-700 nm]
  Visible --> NIR[NIR 700-1000 nm]  
  NIR --> SWIR[SWIR 1.0 to 2.5 µm]
  SWIR --> MWIR[MWIR]
```

::: {.notes}
Spectral remote sensing exploits unique reflectance patterns of leaves and canopies in visible (VIS), near-infrared (NIR) and shortwave infrared (SWIR) bands to assess crop health. Vegetation indices such as NDVI (Normalized Difference Vegetation Index), PRI (Photochemical Reflectance Index) and EVI (Enhanced Vegetation Index) quantify chlorophyll content, photosynthetic efficiency and biomass. Hyperspectral sensors with hundreds of narrow bands enable fine-scale detection of nutrient deficiencies (e.g., nitrogen, phosphorus) and early water stress by identifying specific absorption features. Key steps include sensor radiometric calibration, atmospheric correction, and geometric registration to produce accurate reflectance maps. Integration of UAV-based hyperspectral imaging with satellite platforms allows temporal monitoring at field and regional scales. Finally, spectral data are processed through GIS and decision support systems to generate variable-rate input maps, optimizing fertilizer and irrigation applications to match spatial variability and improve yield and resource-use efficiency.
:::

---

## Visible Region Interaction

- Visible spectrum reflectance indicates canopy chlorophyll absorption characteristics
- Red and green bands used to compute vegetation indices (e.g., VARI, Green NDVI) for biomass and stress detection
- Leaf structure, pigment concentration, and soil background influence reflectance signal
- High-frequency visible imagery captures temporal dynamics critical for dynamic MZ refinement
- Fusion with soil EC, elevation, and yield data produces robust multi-source zone delineation
- Constraints include variable illumination, shadows, and sensor radiometric calibration drift

::: {.notes}
Reflectance in the visible region (400–700 nm) is heavily modulated by chlorophyll absorption peaks in the blue and red bands and by leaf scattering in the green band. Indices such as the Visible Atmospherically Resistant Index (VARI) and variants of Green NDVI leverage these properties to quantify canopy biomass and stress with high sensitivity. When integrated with static soil electrical conductivity (EC) and topographic elevation maps, visible-region metrics improve spatial resolution of management zones by capturing in-season variability. Frequent imaging schedules mitigate temporal aliasing, enabling dynamic recalibration of zone boundaries. However, practitioners must account for illumination geometry, shadowing effects, and sensor calibration consistency to maintain the integrity of reflectance-derived MZ prescriptions.
:::

---

## Near-Infrared (NIR) Region

- Wavelength range: approximately 700–1300 nm
- High reflectance by healthy vegetation; minimal reflectance by soil and water
- Foundation for vegetation indices such as NDVI and SAVI
- Sensitive to chlorophyll content, plant structure, and biomass
- Critical input for PCA-driven dimensionality reduction in MZ clustering
- Improves spectral contrast among management zones for precise delineation

::: {.notes}
The Near-Infrared (NIR) region, spanning roughly 700 to 1300 nm, is a cornerstone in precision agriculture remote sensing. Healthy vegetation exhibits strong NIR reflectance due to internal leaf structure and low pigment absorption, whereas soils and water appear dark, providing clear spectral separation. Common vegetation indices like NDVI and SAVI leverage this contrast to quantify plant vigor, biomass, and stress levels. In our study, NIR bands were integral to the principal component analysis that reduced multivariate sensor data before clustering. By capturing subtle variations in chlorophyll content and canopy architecture, NIR-derived features enhanced the discrimination power of less conventional clustering algorithms — such as McQuitty’s method and Fanny — ultimately yielding management zones with lower within-zone yield variance and higher spatial continuity. When designing MZ delineation workflows, ensuring high-quality NIR acquisition (through appropriate sensor calibration and atmospheric correction) is critical for reliable spectral inputs and robust zone partitioning.
:::

---

## Shortwave Infrared (SWIR) Region

- Spectral range: 1.0–2.5 µm, capturing strong water absorption features
- High sensitivity to canopy water content and plant stress levels
- Improves intra-zone homogeneity assessment and reduces within-zone yield variance


::: {.notes}
The SWIR region (1.0–2.5 µm) is critical in precision agriculture due to its sensitivity to water absorption peaks around 1.4 µm and 1.9 µm. Research shows SWIR reflectance decreases sharply with increasing leaf and soil moisture content, enabling accurate mapping of crop water stress and soil moisture heterogeneity. SWIR data can penetrate sparse canopies, revealing subsurface soil properties that are obscured in visible and NIR bands. Distinct SWIR spectral signatures have been validated for crop species discrimination, aiding in targeted management zone (MZ) delineation. When integrated with advanced clustering algorithms—such as McQuitty’s linkage and Fanny—SWIR-derived features contribute to statistically robust MZs by maximizing inter-zone differentiation and minimizing intra-zone yield variance, as demonstrated in multi-year field trials in Paraná, Brazil.

Next, examine how SWIR indices integrate with other spectral bands for multi-index composite analysis.
:::

---

## Vegetation Indices: Overview

- Definition and calculation of NDVI (Normalized Difference Vegetation Index).
- Extraction of Landsat-derived NDVI time series at key growth stages: tillering, stem elongation, heading, and maturity.
- Geostatistical workflow: semivariogram modeling and simple kriging to interpolate NDVI to a uniform 30 m grid.
- Pearson correlation to quantify NDVI–yield relationships and select top-performing acquisition dates.
- Composite integration of five high-correlation NDVI maps with soil properties (clay content, total limestone, total nitrogen).
- Delineation of management zones showing 83% pixel-scale agreement with historical yield patterns.

::: {.notes}
In this slide we present an end-to-end workflow for using NDVI as a proxy for crop vigor and biomass in precision agriculture. We begin with normalized difference calculations on Landsat multispectral bands to generate time series at critical phenological stages. These time series are then subjected to semivariogram analysis and simple kriging to produce continuous 30 m resolution NDVI surfaces, enabling consistent spatial comparison.

Next, Pearson correlation coefficients are calculated between interpolated NDVI layers and seven years of georeferenced yield data for wheat, maize, sunflower, and sorghum. This step identifies the most yield-predictive dates—typically during stem elongation and heading—guiding selection of the top five NDVI acquisitions.

These NDVI maps are then normalized and fused with grid-sampled soil attributes (clay, limestone, total nitrogen) to form an eight-layer composite. Fuzzy clustering on this dataset yields management zones that align with yield-based zones at an 83% pixel agreement, demonstrating the robustness of buffered multi-year, multi-crop NDVI integration for delineating precision fertilization zones.
:::

---

## NDVI & EVI

- NDVI = (NIR – Red) / (NIR + Red); normalized difference index, range –1 to +1
- EVI = G*(NIR – Red) / (NIR + C1*Red – C2*Blue + L); enhanced for dense vegetation
- NDVI saturates at high LAI; EVI reduces soil-background and atmospheric effects
- EVI uses blue band for aerosol correction; NDVI is simpler but more noise-prone
- Applications: crop vigor monitoring, biomass estimation, stress detection at field scale
- In Smart-Map (Kriging), NDVI/EVI rasters pass Moran’s I filtering and serve as SVM covariates

::: {.notes}
Begin by defining NDVI and its utility in detecting green biomass; explain its limitation due to saturation when leaf area index (LAI) exceeds moderate values. Introduce EVI with Huete et al. (2002) formulation: gain factor (G), canopy background adjustment (L), and aerosol coefficients (C1, C2) improve response in high-biomass conditions. Cite studies showing EVI maintains sensitivity up to LAI of ~6, whereas NDVI plateaus beyond LAI of ~3. Discuss practical use: both indices are calculated from multispectral imagery, then tested for spatial autocorrelation via Moran’s I in Smart-Map. Only the most informative indices are retained for SVM interpolation of soil properties, enhancing R² by up to 0.83 and reducing RMSE substantially. Emphasize that including EVI alongside NDVI can improve model robustness against atmospheric noise and soil background variability.
:::

---

## Additional Vegetation Indices

- Enhanced Vegetation Index (EVI): reduces atmospheric scattering by incorporating the blue band
- Soil Adjusted Vegetation Index (SAVI): corrects soil background effects under sparse canopies
- Green Normalized Difference Vegetation Index (GNDVI): improves chlorophyll sensitivity via green reflectance
- Red-Edge NDVI (NDRE): exploits red-edge reflectance for detailed nitrogen-status monitoring
- Photochemical Reflectance Index (PRI): tracks xanthophyll cycle dynamics as an early stress indicator
- Fusion of multiple indices via kriging enhances spatial precision in management zone delineation

::: {.notes}
Beyond classical NDVI, advanced indices address limitations arising from soil brightness, atmospheric noise, and nutrient variability. EVI’s blue-band correction yields more stable signals under variable aerosol loads. SAVI’s soil adjustment factor is critical for interpreting low-biomass stages, particularly in early sorghum stands. GNDVI leverages green reflectance to sharpen chlorophyll estimation, correlating strongly with biomass in wheat and maize trials. NDRE, derived from the red-edge region, provides sensitive monitoring of canopy nitrogen status, guiding variable-rate fertilization. PRI reflects changes in the xanthophyll cycle, offering an early diagnostic of plant physiological stress. When these indices are interpolated at 30 m resolution by simple kriging and integrated into fuzzy c-means clustering, the resulting management zones exhibit higher agronomic relevance and maintain strong spatial agreement (≥ 83 %) with multi-year yield maps, thus improving site-specific prescription accuracy.
:::

---

## Time Series Data: Crop Development

- High-frequency temporal sampling for phenological stage monitoring (daily to weekly intervals)
- NDVI/EVI trajectories from satellite and UAV imagery to assess canopy vigor
- Thermal time series for early detection of water stress and canopy temperature dynamics
- Soil moisture and nutrient probe data streams for sub-surface water and fertility tracking
- Growth rate curve fitting and time-series modeling for yield prediction and anomaly detection
- Multi-sensor data fusion: combining satellite, aerial, and ground sensors into unified timelines

::: {.notes}
Time series data are foundational for precision crop management because they capture dynamic changes in plant development, water status, and nutrient uptake. High-frequency sampling—from daily satellite overpasses to sub-daily UAV flights and ground sensor recordings—enables accurate phenological staging (e.g., emergence, tillering, heading) and supports real-time decision making. Vegetation indices like NDVI and EVI plotted over time reveal canopy growth curves, while thermal time series highlight early water stress before visible wilting. Soil moisture and nutrient probe outputs provide complementary sub-surface data, linking above-ground observations to root-zone conditions. Fitting growth rate models to these data streams allows prediction of final yield and identification of growth anomalies (e.g., pest outbreaks or nutrient deficiencies). Finally, fusing multi-source time series through geostatistical interpolation and machine-learning frameworks enhances spatial-temporal resolution, reduces uncertainty, and underpins variable-rate application algorithms for optimized input use.
:::

---

## Stress Detection Over Time

:::: {.columns}

::: {.column}

```{mermaid}
graph TD
  A[NDVI Time Series] -->|Detects Anomalies| D[Stress Event]
  B[Canopy Temperature] -->|Heat Stress Indication| D
  C[Soil Moisture Logs] -->|Drought Signal| D
  D --> E[Update Management Zone]
  E --> F[Variable-Rate Response]
```
:::

::: {.column}
- Multi-temporal NDVI analysis reveals onset, peak, and recovery phases of plant stress
- Canopy temperature and vapor pressure deficit (VPD) time series identify heat stress events
- Soil moisture sensor logs detect drought stress duration and severity
- Spatio-temporal maps within management zones (MZs) localize stress hotspots
- Bayesian spatio-temporal modeling integrates multi-year data for stress forecasting
- Fusion of satellite, aerial, and ground sensors enhances temporal resolution and reliability

:::
::::

::: {.notes}
This slide highlights the methods used to detect and monitor crop stress over time in precision agriculture. Multi-temporal NDVI trajectories allow us to pinpoint when vegetation deviates from expected growth curves, indicating stress onset and recovery. Canopy temperature coupled with vapor pressure deficit provides early warnings of heat stress by detecting changes in plant transpiration dynamics. Soil moisture time series from in-field sensors quantify drought periods and help differentiate moisture-related stress from other factors. By mapping these signals within predefined management zones (MZs), we can spatially localize stress hotspots and adapt interventions with fine resolution. Advanced Bayesian spatio-temporal models leverage multi-year datasets to forecast stress patterns and optimize nutritional or irrigation responses. Finally, fusing data from satellites, drones, and proximal sensors increases temporal revisit rates and improves detection accuracy under variable environmental conditions. This comprehensive approach enables proactive stress management, reducing yield losses and input waste.
:::

---

## Yield Prediction Using Time Series

- Integrate sequential crop sensor metrics (e.g., NDVI, canopy temperature) over key growth stages
- Apply ARIMA and LSTM time-series models for capturing temporal yield dynamics and anomalies
- Incorporate soil texture and structure covariates to account for spatial heterogeneity
- Produce high-resolution yield maps with prediction intervals at zonal and whole-field scales
- Validate model outputs against harvest monitor data using cross-seasonal cross-validation
- Leverage forecasts for zone-specific input optimization in Site-Specific Crop Management (SSCM)

::: {.notes}
Yield prediction via time-series analysis combines remote sensing metrics and environmental covariates to model temporal crop performance. ARIMA models capture linear trends and seasonality, while LSTM neural networks handle nonlinear dependencies and long-term memory of growth patterns. Adding soil texture (sand, silt, clay proportions) and soil structure (pore connectivity, compaction status) as exogenous inputs ensures that spatial variability in nutrient and moisture dynamics is embedded in the forecasts. We generate continuous yield maps with uncertainty bounds, enabling agronomists to identify under-performing zones and calibrate inputs accordingly. Model validation relies on aligning predicted yields with actual harvest data through cross-seasonal and k-fold validation schemes, providing robust accuracy estimates. The resulting time-series forecasts inform SSCM by prescribing differential fertilizer, irrigation, and seeding rates tailored to temporal and spatial yield potential, optimizing resource use and maximizing productivity.
:::

---

## Blackbody Radiance

```{r}
library(ggplot2)
library(ggplot2)

# Constants
h <- 6.62607015e-34  # Planck's constant (J·s)
c <- 2.99792458e8    # Speed of light (m/s)
k <- 1.380649e-23    # Boltzmann constant (J/K)

# Planck's law function for spectral radiance
planck <- function(lambda, T) {
  (2 * h * c^2 / lambda^5) / (exp(h * c / (lambda * k * T)) - 1)
}

# Wavelengths (in meters, 300nm to 2500nm)
lambda <- seq(300e-9, 2500e-9, length.out = 1000)

# Temperatures (in Kelvin, including Sun at ~5778K)
temps <- c(3000, 4000, 5000, 5778)
temp_labels <- c("3000K", "4000K", "5000K", "Sun (5778K)")

# Create data frame
data <- data.frame()
for (T in temps) {
  radiance <- planck(lambda, T)
  temp_data <- data.frame(Wavelength = lambda * 1e9, Radiance = radiance, Temperature = as.factor(T))
  data <- rbind(data, temp_data)
}

# Plot
ggplot(data, aes(x = Wavelength, y = Radiance, color = Temperature)) +
  geom_line() +
  scale_color_manual(values = c("red", "orange", "green", "yellow", "blue"), labels = temp_labels) +
  labs(title = "Blackbody Radiance Curves (Wien's Law)",
       x = "Wavelength (nm)",
       y = "Spectral Radiance (W/m²/sr/m)") +
  theme_minimal()
```





---

## Vegetation Spectra

```{r}
library(prosail)
# define input variables for PROSPECT. 
# refer to prospect tutorial for default values
Input_PROSPECT <- data.frame('CHL' = 40, 'CAR' = 8, 'ANT' = 0.0, 
                             'EWT' = 0.01, 'LMA' = 0.009, 'N' = 1.5)
# define input variables for SAIL. 
lai <- 5      # LAI
q <- 0.1      # Hot spot parameter
TypeLidf <- 2 # leaf inclination distribution function 
LIDFa <- 30   # mean leaf angle
tts <- 30     # geometry of acquisition: sun zenith angle
tto <- 10     # geometry of acquisition: observer zenith angle
psi <- 90     # geometry of acquisition: sun-observer azimuth 
rsoil <- SpecSOIL$Dry_Soil # soil reflectance (SpecSOIL includes Dry_Soil and Wet_Soil properties)

# run PROSAIL with 4SAIL
Ref_4SAIL <- PRO4SAIL(Input_PROSPECT = Input_PROSPECT, 
                      TypeLidf = TypeLidf, LIDFa = LIDFa, lai = lai,
                      q = q, tts = tts, tto = tto, psi = psi, rsoil = rsoil)

# run PROSAIL with 4SAIL2
# define leaf chemical constituents corresponding green and brown leaves
Input_PROSPECT <- data.frame('CHL' = c(40,5), 'CAR' = c(8,5), 'ANT' = c(0,1), 
                             'EWT' = c(0.01,0.005), 'LMA' = c(0.009,0.008), 
                             'BROWN' = c(0.0,0.5), 'N' = c(1.5,2))
# define additional 4SAIL2 parameters
fraction_brown <- 0.5
diss <- 0.5
Cv <- 1
Zeta <- 1

# run PROSAIL with 4SAIL2
Ref_4SAIL2 <- PRO4SAIL(SAILversion = '4SAIL2',
                       Input_PROSPECT = Input_PROSPECT,
                       TypeLidf = TypeLidf, LIDFa = LIDFa, lai = lai,
                       q = q, tts = tts, tto = tto, psi = psi, rsoil = rsoil,
                       fraction_brown = fraction_brown, diss = diss, 
                       Cv = Cv, Zeta = Zeta)
#ggplot the Ref_4SAIL2 relectance  rsot
data <- data.frame(Wavelength = c(400:2500), 
                   Reflectance = Ref_4SAIL2$rdot)
# Convert wavelength to nanometers
data$Wavelength <- data$Wavelength
# Convert reflectance to percentage
data$Reflectance <- data$Reflectance*100 
# Rename columns for clarity
colnames(data) <- c("Wavelength", "Reflectance")
# ggplot it




# Plot reflectance
library(ggplot2)
ggplot(data, aes(x = Wavelength, y = Reflectance)) +
  geom_line(color = "darkgreen") +
  labs(title = "Plant Canopy Reflectance (PROSAIL Simulation)",
       x = "Wavelength (nm)",
       y = "Reflectance") +
  theme_minimal()
```
---

## Sentinel Convolution

```{r}

 
  #get the spectral response for Sentinel-2A
  SensorName <- 'Sentinel_2A'
  # if interested in a different satellite, please use Path_SensorResponse to 
  # locate the SRF file expected to be named 'SensorName_Spectral_Response.csv' 
  # (separator = tabulations)
  SRF_S2 <- GetRadiometry(SensorName,Path_SensorResponse = NULL)
  WL <- seq(400, 2400, by = 10)
  # FWHM is a vector giving FWHM for each band in nanometer
  FWHM <- 10+ 0*seq(400, 2400, by = 10)
  # compute corresponding SRF
  SRF_SH <- Compute_SRF(WL,FWHM)
  
  wvl <- SpecPROSPECT_FullRange$lambda
  BRF_4SAIL_S2 <- applySensorCharacteristics(wvl = wvl,
                                             SRF = SRF_S2,
                                             InRefl = Ref_4SAIL2)

  
  ##ggplot BRF_4SAIL_S2 reflectance
data_s2 <- data.frame(Wavelength = SRF_S2$Central_WL, Reflectance = BRF_4SAIL_S2$rdot)
# Convert reflectance to percentage
data_s2$Reflectance <- data_s2$Reflectance * 100
# Rename columns for clarity
colnames(data_s2) <- c("Wavelength", "Reflectance")
# Plot reflectance for Sentinel-2A
ggplot(data_s2, aes(x = Wavelength, y = Reflectance)) +
  geom_line(color = "blue") +
  labs(title = "Sentinel-2A Canopy Reflectance (PROSAIL Simulation)",
       x = "Wavelength (nm)",
       y = "Reflectance") +
  theme_minimal()
```

---

## Spectra by LAI

```{r}
run_prosail_lai <- function(lai) {
  Input_PROSPECT <- data.frame('CHL' = 40, 'CAR' = 8, 'ANT' = 0.0, 
                               'EWT' = 0.01, 'LMA' = 0.009, 'N' = 1.5)
  
  Ref_4SAIL <- PRO4SAIL(Input_PROSPECT = Input_PROSPECT, 
                        TypeLidf = TypeLidf, LIDFa = LIDFa, lai = lai,
                        q = q, tts = tts, tto = tto, psi = psi, rsoil = rsoil)
  
  return(Ref_4SAIL$rdot)
}
# Define a range of LAI values
lai_values <- seq(1, 10, by = 1)
# Run PROSAIL for each LAI value and store results
results <- lapply(lai_values, run_prosail_lai)
# Create a data frame for plotting
data_lai <- data.frame(Wavelength = rep(SpecPROSPECT_FullRange$lambda, length(lai_values)),
                       Reflectance = unlist(results),
                       LAI = rep(lai_values, each = length(SpecPROSPECT_FullRange$lambda)))
# Plot the results using ggplot2
ggplot(data_lai, aes(x = Wavelength, y = Reflectance, color = as.factor(LAI))) +
  geom_line() +
  labs(title = "Canopy Reflectance with Varying LAI",
       x = "Wavelength (nm)",
       y = "Reflectance",
       color = "Leaf Area Index (LAI)") +
  theme_minimal() +
  scale_color_brewer(palette = "Set1")  
```

---

## Spectral Resolution Degradations

```{r}
data_multispectral <- data.frame(Wavelength = data_s2$Wavelength,
                                  Reflectance = data_s2$Reflectance,
                                  Type = "Multispectral (Sentinel-2A)")
data_hyperspectral <- data.frame(Wavelength = data$Wavelength,
                                  Reflectance = data$Reflectance,
                                  Type = "Hyperspectral (PROSAIL Simulation)")
# Combine the two data frames
data_combined <- rbind(data_multispectral, data_hyperspectral)
# Plot the combined data using ggplot2
ggplot(data_combined, aes(x = Wavelength, y = Reflectance, color = Type)) +
  geom_line() +
  labs(title = "Multispectral vs Hyperspectral Reflectance",
       x = "Wavelength (nm)",
       y = "Reflectance") +
  #legend on the bottom 
  theme_minimal() +
  theme(legend.position = "bottom") +
  scale_color_manual(values = c("blue", "darkgreen"))
```

---

## Radiance Ground and Vegetation 

```{r}
library(photobiology)

data(sun.spct)  # Example solar spectrum at ground level


#for TOA irradiance you can use the planck function with a temperature of 5778K
# Multiply the ground radiance by reflectance to get the radiance reflected by the vegetatation. The information an be founnd on SRF_SH$Spectrl_Response
# Create data frame for plotting
data <- data.frame(
  Wavelength = sun.spct$w.length[120:522],
  Ground_Radiance = sun.spct$s.e.irrad[120:522],
  Vegetation_Radiance = t((sun.spct$s.e.irrad[120:522]) * t(results[[1]][0:403])) # Example vegetation reflectance
#  TOA_Radiance = planck(sun.spct$w.length * 1e-9, 5778) * sun.spct$s.e.irrad / (h * c)
)
    
# Plot irradiance at ground level
ggplot(data, aes(x = Wavelength)) +
  geom_line(aes(y = Ground_Radiance, color = "Ground Level"), linewidth = 1) +
 geom_line(aes(y = Vegetation_Radiance, color = "Radiance Vegetation"), linewidth = 1, linetype = "dashed") +
  labs(title = "Irradiance at Ground vs. Radiance Vegetation ",
       x = "Wavelength (nm)",
       y = "Spectral Irradiance (W/m²/nm)") +
 # scale_color_manual(values = c("Ground Level" = "blue", "Top of Atmosphere" = "red")) +
  theme_minimal() + theme(legend.position = "bottom") 
```

---

## Spectral Data & Remote Sensing Basics

- EM and spectral sensors capture subsurface conductivity profiles and canopy reflectance for soil moisture and crop stress mapping
- Multispectral sensors record visible (400–700 nm) and near-infrared (700–1000 nm) bands at high spatial resolution (cm-level via UAVs)
- Vegetation indices (e.g., NDVI, NDRE, MSAVI) quantify biomass, chlorophyll content, and water status
- Hyperspectral imaging (400–2500 nm) provides ~10 nm spectral resolution to detect specific pigment and water absorption features
- Radiometric calibration and atmospheric correction routines convert raw radiance to quantitative surface reflectance
- Ground-truthing with geolocated soil cores and neutron moisture probes calibrates spectral moisture proxies and validates spatial models

::: {.notes}
Spectral remote sensing provides non-invasive, spatially continuous measurements of canopy and soil properties crucial for precision agriculture. Multispectral sensors—commonly mounted on UAV or satellite platforms—capture broad VIS and NIR bands used to compute vegetation indices that correlate strongly with biomass, chlorophyll, and water content. Hyperspectral systems extend this capability by offering narrow spectral bands (∼10 nm) across 400–2500 nm, enabling discrimination of subtle pigment variations and water absorption features. To ensure quantitative comparability, raw sensor data undergo radiometric calibration against known reflectance panels and atmospheric correction to remove path radiance and aerosol effects. Calibration with ground-truth data—such as volumetric water content from soil cores and neutron moisture meter measurements—facilitates regression models that translate reflectance metrics into accurate soil moisture and crop stress estimates. This slide lays the foundation for integrating spectral and EM datasets to enhance field-scale decision support.

Next, we will examine methods for fusing spectral and EM data to build predictive soil moisture models.
:::

---



## Remote Sensing in Agriculture

- Evolution from discrete soil-sampling grids to continuous sensor-based spectral data acquisition
- Platforms: satellite (Landsat, Sentinel) vs. UAVs with multispectral, hyperspectral, thermal, and LiDAR sensors
- Vegetation indices (NDVI, NDRE, MSAVI) for quantifying crop vigor and nutrient status
- Thermal infrared imaging for early water stress detection; LiDAR for canopy structure and biomass estimation
- Data processing: radiometric correction, orthorectification, spectral index computation, and classification workflows
- Integration with GIS and geostatistics (kriging) to generate high-resolution variable-rate application maps

::: {.notes}
Remote sensing has transformed precision agriculture by providing spatially continuous data across entire fields, overcoming the spatial gaps inherent in traditional soil-sampling grids pioneered in the 1960s and refined via kriging in the 1980s. Satellite platforms such as Landsat and Sentinel deliver 10–30 m resolution imagery with multi-day revisit intervals, suitable for broad monitoring. UAVs equipped with multispectral, hyperspectral, thermal, and LiDAR sensors enable centimeter-level resolution, capturing fine-scale variability in crop health and soil properties.

Vegetation indices—NDVI, NDRE, MSAVI—leverage red and near-infrared reflectance to assess biomass, chlorophyll, and nitrogen status, while thermal infrared imaging identifies canopy temperature anomalies indicative of water stress. LiDAR provides 3D canopy structure data for accurate biomass estimation. Collected data undergo radiometric calibration, atmospheric correction, orthorectification, and spectral index calculation before GIS integration.

Combined with geostatistical interpolation methods (e.g., kriging), these datasets form the basis for prescription maps used in variable-rate nutrient and irrigation applications. Challenges remain in sensor calibration, atmospheric effects, data volume management, and multi-source data fusion, but the foundational geostatistical work of Mulla and colleagues enables today’s high-resolution, cost-effective site-specific management practices.
:::


---

## Spatial Patterns from Pathogens

- Soil electrical conductivity (EC), resistivity, organic matter, texture, and depth mapping for static zone delineation.
- Vegetation indices (e.g., NDVI) from satellite, aerial, and proximal sensors capturing dynamic canopy chlorophyll and biomass patterns.
- In-season remote sensing integration refines temporal variability over static soil-based maps.
- Hybrid clustering frameworks combining EC, elevation, NDVI, canopy volume, and yield to define nutrient, irrigation, and pest management zones.
- Spatio-temporal Bayesian models integrating multi-source covariates for predictive mapping of nitrogen requirements and yield variability.
- Case studies: citrus productivity zones, wheat nitrogen management, and irrigation zoning via soil water capacity and multispectral imagery.

::: {.notes}
Precision agriculture relies on accurate delineation of Management Zones (MZs) to apply variable-rate inputs efficiently. Static soil property mapping using EC, resistivity, organic matter, texture, and depth measurements provides baseline spatial structure but often fails to capture temporal yield variability alone.

In-season crop-based sensing, including NDVI from satellite, aerial, or proximal platforms, quantifies canopy chlorophyll, biomass, and leaf area to detect crop stress and productivity changes. Integrating these dynamic observations with soil data enhances zoning precision, capturing both stable and transient patterns.

Hybrid approaches leverage clustering algorithms to combine multi-source covariates—soil EC, elevation, NDVI, canopy volume, and historical yield records—into agronomically relevant MZs for nutrients, irrigation, and pest management.

Advanced spatio-temporal Bayesian modeling further refines zone delineation by probabilistically linking EC and crop trial data with observed yield responses over time, enabling predictive nitrogen mapping under uncertainty. Illustrative studies include citrus orchards partitioned into productivity zones via EC, NDVI, and canopy volume; wheat fields modeled for nitrogen zones; and irrigation zones derived from soil water holding capacity coupled with imagery and yield data. These integrated frameworks drive site-specific management, optimizing input use and enhancing sustainability.
:::



## Abiotic Stress Patterns

- Spatial mapping of soil moisture and nutrient variability to delineate abiotic stress hotspots
- Remote sensing indices (NDVI, thermal IR) for early detection of drought and heat stress
- Variable-rate irrigation and fertilization based on stress zone boundaries and temporal crop demand
- Conservation structures (grassed waterways, contour furrows) to reduce erosion under extreme precipitation events
- Precision residue harvesting and soil-specific tillage to improve moisture retention and mitigate compaction
- Riparian buffers and constructed wetlands for off-field interception of nutrient and sediment runoff

::: {.notes}
Abiotic stress patterns in agricultural fields manifest as spatial–temporal variability in water availability, nutrient distribution, temperature extremes, and erosion risk. High-resolution soil moisture sensors and geospatial mapping pinpoint stress hotspots with meter-scale accuracy. Satellite and drone-based NDVI and thermal infrared imagery provide continuous monitoring, enabling early detection of water deficit and heat stress before crop wilting or yield loss occurs. Variable-rate irrigation and nutrient application technologies align resource delivery with real-time crop demand and stress thresholds, potentially boosting nutrient use efficiency by 20–30%.

Conservation structures such as grassed waterways and contour furrows are vital under climate-change scenarios that project a 1.7% increase in soil erosion per 1% rise in precipitation. Precision residue management and soil-specific tillage further enhance soil structure and moisture retention, alleviating compaction and runoff. Off-field practices, including riparian buffers and constructed wetlands, intercept nutrient and sediment flows at the watershed scale, closing the loop on nutrient losses. Integrating these approaches builds resilient cropping systems capable of maintaining high yields amid evolving abiotic challenges and shrinking water resources.
:::

---


## Routine for Crop Scouting

- Establish grid-based scouting paths (30–60 m spacing)
- Use GPS-enabled devices for georeferenced field data
- Monitor pest, disease, and nutrient status with proximal sensors and UAV imagery
- Record soil moisture and temperature using in-field probes
- Integrate NDVI and thermal imaging for early stress detection
- Upload and sync data to GIS for spatial analysis and intervention mapping

```{mermaid}
graph LR
  A[Plan Grid Sampling Paths] --> B[Field Data Collection]
  B --> C[Data Upload & QC]
  C --> D[Geostatistical Analysis]
  D --> E[Site-Specific Interventions]
  E --> F[Review & Adjust]
```

::: {.notes}
Crop scouting routines in precision agriculture leverage grid-based sampling at recommended spacings of 30–60 m to ensure sufficient spatial resolution for geostatistical modeling. Utilizing GPS-enabled handhelds or tablets, agronomists record pest pressure, disease incidence, and plant nutrient status directly in the field, ensuring precise geolocation. Proximal sensors and UAV-mounted multispectral/thermal cameras augment traditional observations by providing high-resolution NDVI and canopy temperature data, enabling early stress detection. Data are uploaded to a GIS platform where kriging or inverse-distance weighting creates contour maps of key biophysical variables, guiding variable-rate fertilizer, pesticide, and irrigation applications. This workflow supports a closed-loop management system, optimizing inputs, reducing environmental impact, and improving yield consistency across heterogeneous fields.
:::

---

## Monitoring & Data Logging Routine

- Pre-start of cropping season (ECa soil clusters)
- In-season NDVI acquisition from UAV and satellite platforms for canopy chlorophyll mapping (target scouting)
- Harvest Yield Maps (end-results)
- Integration into spatio-temporal database enabling dynamic MZ delineation and trend analysis

::: {.notes}
Accurate monitoring and rigorous data logging form the backbone of precision agriculture management zones (MZs). We deploy soil probes measuring EC and moisture at high sampling frequencies (e.g., 30-minute intervals) to capture diurnal and irrigation-driven dynamics. Each probe is geo-tagged using RTK-GNSS, ensuring sub-decimeter placement accuracy critical for delineating small, homogeneous MZs.

Simultaneously, we schedule weekly NDVI flights or satellite passes to quantify canopy chlorophyll and biomass patterns, which often precede yield variability. Data loggers at field gateways buffer high-resolution streams locally, then push to a cloud-based GIS repository to prevent loss during network outages.

Our preprocessing pipeline applies statistical QC: flagging sensor drifts, removing outliers, infilling missing records through spatio-temporal kriging. The cleaned datasets feed into a central spatio-temporal database, where MZ boundaries are dynamically updated based on combined soil and plant metrics. This routine supports case studies such as citrus nutrient zoning and wheat N-management, and underpins variable-rate irrigation, fertilization, and chemical applications.
:::


